---
title: "Pressure"
author: "Alex Vand"
date: "`r Sys.Date`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "Final Project"
subtitle: "EDS 222"
author: "Alex Vand"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install packages

```{r, warning = FALSE}
library(tidyverse)
library(tsibble)
library(zoo)
library(feasts)
library(janitor)
library(dplyr)
```


## Read in the raw data (.csv from NOAA)
- https://www.cpc.ncep.noaa.gov/data/indices/soi

Tahiti-Darwin sea level press(ure?) anomaly

```{r}
data_raw_pressure <- read_csv(here::here("pressure.csv"))

data_pressure <- janitor::clean_names(data_raw_pressure)
#                          pivot_longer(data_pressure, jan:dec)
```

```{r}
data_pressure %>% 
  pivot_longer(!"year", names_to = "month", values_to = "anom")
```




## Lag analysis

Statistical lag analysis is implemented through the autocorrelation function `acf`, where the argument `lax.max = ` is indicated/specified in each of the following scenarios:
- 478

$t-125$

$t-175$

$t-390$

```{r}
acf4 <- acf(x = data$anom_4, lag.max = 478)
```



```{r}
acf6 <- acf(x = data$anom_6, lag.max = 478)
```



```{r}
acf8 <- acf(x = data$anom_8, lag.max = 478)
```



```{r}
acf10 <- acf(x = data$anom_10, lag.max = 478)
```



```{r}
acf10 <- acf(x = data$anom_10, lag.max = 200)
```



```{r}
max(acf10$lag[acf10=0.2])

sort(acf10$lag[acf10=-0.3])
```


## Autocorrelation

ADL?

```{r}

```


## Hypothesis testing?

```{r}
data_mean <- data_raw %>% group_by(MON) %>%
         summarize(avg_anom4 = mean(ANOM...4),
                   avg_anom6 = mean(ANOM...6),
                   avg_anom8 = mean(ANOM...8),
                   avg_anom10 = mean(ANOM...10))

data_sd <- data_raw %>% group_by(MON) %>%
         summarize(sd_anom4 = sd(ANOM...4),
                   sd_anom6 = sd(ANOM...6),
                   sd_anom8 = sd(ANOM...8),
                   sd_anom10 = sd(ANOM...10))
```


## Simple linear regression

```{r}
model <- summary(lm(YR ~ ANOM...4, data = data_raw))
model
```

```{r}
model <- summary(lm(MON ~ avg_anom4, data = data_mean))
model
```


## Classical decomposition

```{r}
data <- data %>% 
  mutate(month = case_when(
    mon <= 9 ~ paste0(0, mon),
    mon > 9 ~ paste0(mon))) %>%
  unite(col = date, yr, month, sep = "-") %>% 
  mutate(date = yearmonth(as.yearmon(date, format = '%Y-%m'))) %>%
  na.omit() %>%
  select(date, anom_10) %>% 
  as_tsibble(index = date)
```


```{r}
data %>% model(classical_decomposition(anom_10, type = "additive")) %>%
  components() %>%
  autoplot() +
  labs(title = "Classical additive decomposition of ENSO SST anomalies")
```

https://tsibble.tidyverts.org/reference/as-tsibble.html

## Question
## Background and Motivation
## Data
## Analysis and Results
## Future Research
Future research aims to hopes to 


```{r}
data_raw_cassava <- read_csv(here::here("cassava_yields.csv"))

data_cassava <- janitor::clean_names(data_raw_cassava)
```

```{r}
as_tsibble(data_cassava, index = year)
```


```{r}
data_cassava %>% model(classical_decomposition(crops_cassava_125_yield_5419_hg_ha, type = "additive")) %>%
  components() %>%
  autoplot() +
  labs(title = "Classical additive decomposition of ENSO SST anomalies")
```



